A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-03-18 paper/draft.tex
generate partisannship spells

Note that party affiliation in Brazil is high, with over 10 percent of nationally registered voters affiliated to a political party \citep{speck2015estudo}. Partisan ties at the electorate level are stable, in contrast with more strategic party switching occurs at the political candidate level \citep{desposato2006parties}. As of 2019, the year for which party affiliation data was extracted, there were over 11 million active party affiliates, with around 5 million registries either canceled or terminated by party members.

Additionally, it is worthwhile noting that the affiliation of party members has been increasing over time. Importantly, this evolution has been rather rapid over time. While in the year 1997 there were around 200 thousand active party members, by the year 2018 this has increased to 10 million.

@@

\section{Descriptive statistics}
\label{sec:descriptive}

In this section, I provide descriptive statistics that illustrate broader patterns in the distribution of party affiliated bureaucrats at the municipal level in Brazil. Figure \ref{fig:map_pooled} illustrates the subnational distribution of party affiliation in Brazil for the year of 2015. There are a few patterns that emerge. The first one is that there is

First, that the distribution of partisan affiliation varies significantly over the national territory, with its predominance in the Northeast, as well as in the Northern regions of the country. The Midwest and Southeast appear to have the lowest prevalence of partisan affiliation. Second, that there is wide variation with respect to the prevalence of partisanship. In particular, some municipalities are almost entirely occupied by party members, with over 60 percent of its workforce affiliated to a party. Others are largely autonomous from party members, with less than 10 percent of municipal workers affiliated to a party.

\begin{figure}[H]
    \centering
    \includegraphics[width = 8cm, height = 8cm]{figures/maps/pooled.png}
    \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}. Darker colors indicate a larger degree of partisan affiliation. This visualization includes all bureaucrats, from high-level managers to service staff.}
    \label{fig:map_pooled}
\end{figure}

>>>>

Note that party affiliation in Brazil is relatively high, with over 10 percent of nationally registered voters affiliated to a political party, as compared with other advanced democracies that hover around 5 percent \citep{speck2015estudo}. Partisan ties at the voter level are stable, in contrast with more strategic party switching occurs at the political candidate level \citep{desposato2006parties}. As of 2019, the year for which party affiliation data was extracted, there were over 11 million active party affiliates, with around 5 million registries either canceled or terminated by party members.

It is important to note that the affiliation of party members has been increasing over time. Importantly, this evolution has been rather rapid over time. While in the year 1997 there were around 200 thousand active party members, by the year 2018 this has increased to 10 million. While the causes of this rapid increase are not clear, it is most likely due to both increased registration and better quality data on partisan affiliation as the collection of information by the national electoral authority.

@@

\subsection{Prevalence of Patronage}
\label{sec:prevalence_patronage}

The distribution of partisan affiliation varies significantly over the national territory, with its predominance in the Northeast, as well as in the Northern regions of the country. The Midwest and Southeast appear to have the lowest prevalence of partisan affiliation. Second, that there is wide variation with respect to the prevalence of partisanship. In particular, some municipalities are almost entirely occupied by party members, with over 60 percent of its workforce affiliated to a party. Others are largely autonomous from party members, with less than 10 percent of municipal workers affiliated to a party.

\begin{figure}[H]
    \centering
    \includegraphics[width = 9cm, height = 9cm]{figures/maps/pooled.png}
    \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}. Darker colors indicate a larger degree of partisan affiliation. Sample includes all bureaucrats, from high-level managers to service staff.}
    \label{fig:map_pooled}
\end{figure}
